Observed vs. Simulated Year Comparison

### 2021-12-11 03-post-process/compare-sim-to-obs.Rmd
Include functions create tables summarizing sim vs. obs ratios

>>>>

```{r summarize_ratios fn}
summarize_ratios = function(x, Class, Type) {
  xsumm = as.data.frame(t(apply(x[,"sim-v-obs",,], 3, colMeans)))
  xsumm = cbind(Class = Class, Type = Type, ID = rownames(xsumm), xsumm); rownames(xsumm) = NULL
  return(xsumm)
}
```

```{r assign_color fns}
# functions to assign a color based on a numeric value

# values close to 1 will be closer to white; values closer to 0 will be blue; values closer to 2 will be red
assign_color_1 = function(x) {
  red_pal = colorRampPalette(c("white", "red"))
  blue_pal = colorRampPalette(c("blue", "white"))
  red_cols = red_pal(5)
  blue_cols = blue_pal(5)
  cols = c(blue_cols, red_cols)

  breaks = c(0, 0.25, 0.5, 0.75, 0.9, 1.1, 1.25, 1.5, 1.75, 3)
  bin_names = levels(cut(runif(1e6, 0, 2), breaks = breaks, include.lowest = TRUE))
  names(cols) = c(bin_names[1:5], bin_names[5:9])
  bins = cut(x, breaks = breaks, include.lowest = TRUE)
  cols_use = cols[as.character(bins)]
  cols_use[is.na(cols_use)] = cols[5]
  alpha(unname(cols_use), 0.5)
}

# values close to 0 will be closer to white; values closer to -1 will be blue; values closer to 1 will be red
assign_color_0 = function(x) {
  red_pal = colorRampPalette(c("white", "red"))
  blue_pal = colorRampPalette(c("blue", "white"))
  red_cols = red_pal(5)
  blue_cols = blue_pal(5)
  cols = c(blue_cols, red_cols)

  breaks = c(-1, -0.75, -0.5, -0.25, -0.1, 0.1, 0.25, 0.5, 0.75, 1)
  bin_names = levels(cut(runif(1e6, -1, 1), breaks = breaks, include.lowest = TRUE))
  names(cols) = c(bin_names[1:5], bin_names[5:9])
  bins = cut(x, breaks = breaks, include.lowest = TRUE)
  cols_use = cols[as.character(bins)]
  cols_use[is.na(cols_use)] = cols[5]
  alpha(unname(cols_use), 0.5)
}
```

```{r prep_est_ratios fn}
# function to prepare ratios for table-making
prep_est_ratios = function(x, q) {
  # add/format additional column IDs
  x$origin = stringr::str_extract(x$ID, "NOR|HOR")
  x$pop = stringr::str_extract(x$ID, "CAT|LOS|MIN|UGR")
  x$LH_type = stringr::str_extract(x$ID, "fall|spring")
  x$LH_type[x$LH_type == "spring"] = "Spring"
  x$LH_type[x$LH_type == "fall"] = "Fall"
  x$pop = ifelse(is.na(x$pop), "Combined", x$pop)
  x$origin = ifelse(is.na(x$origin), "Combined", x$origin)
  x$LH_type = ifelse(is.na(x$LH_type), "Combined", x$LH_type)

  # make populations separate columns
  x = dcast(x, Class + Type + origin + LH_type ~ pop, value.var = q)

  # reorder columns
  x = x[,c("Class", "Type", "LH_type", "origin", "CAT", "LOS", "MIN", "UGR", "Combined")]

  # reorder rows
  x$Class = factor(x$Class, levels = c("FW Juvenile", "SW Juvenile", "FW Adult"))
  x$Type = factor(x$Type, levels = c("Egg to Parr Survival", "Proportion Fall Migrants", "Overwinter Survival", "Migration to LGR Survival", "LGR to Ocean Survival", "First Year Ocean Survival", "Age-3 Maturity", "Age-4 Maturity", "BON to LGR Survival", "Pre-Spawn Survival"))
  x$LH_type = factor(x$LH_type, levels = c("Combined", "Fall", "Spring"))
  x$origin = factor(x$origin, levels = c("Combined", "NOR", "HOR"))
  x = x[order(x$Class, x$Type, x$LH_type, x$origin),]

  colnames(x)[which(colnames(x) == "LH_type")] = "Mig. Type"
  colnames(x)[which(colnames(x) == "Origin")] = "Origin"

  return(x)
}
```

```{r prep_derived_ratios fn}
# function to prepare ratios for table-making
prep_derived_ratios = function(x, q) {
  # add/format additional column IDs
  x$origin = stringr::str_extract(x$ID, "NOR|HOR")
  x$pop = stringr::str_extract(x$ID, "CAT|LOS|MIN|UGR")
  x$LH_type = stringr::str_extract(x$ID, "fall|spring")
  x$LH_type[x$LH_type == "spring"] = "Spring"
  x$LH_type[x$LH_type == "fall"] = "Fall"
  x$pop = ifelse(is.na(x$pop), "Combined", x$pop)
  x$origin = ifelse(is.na(x$origin), "Combined", x$origin)
  x$LH_type = ifelse(is.na(x$LH_type), "Combined", x$LH_type)

  # make populations separate columns
  x = dcast(x, Class + Type + origin + LH_type ~ pop, value.var = q)

  # reorder columns
  x = x[,c("Class", "Type", "LH_type", "origin", "CAT", "LOS", "MIN", "UGR", "Combined")]

  # reorder rows
  x$Class = factor(x$Class, levels = c("Abundance", "Production Rates", "SAR", "Adult Composition"))
  x$Type = factor(x$Type, levels = c("Total Adults (Estuary)", "Total Adults (To Tributary)", "Total Spawners", "Total Egg Production", "Total Parr Recruitment", "Total LGR Smolt", "Aggregate Total Adult Return (To Tributary)", "Aggregate Total Smolt (At LGR)", "Parr per Spawner", "Smolt in Tributary per Spawner", "Spawners per Spawner", "Smolt in Tributary per Parr", "Smolt at LGR per Parr", "BON to BON", "LGR to LGR", "Proportion Hatchery Return", "Proportion Age-3 Return", "Proportion Age-4 Return", "Proportion Age-5 Return"))
  x$LH_type = factor(x$LH_type, levels = c("Combined", "Fall", "Spring"))
  x$origin = factor(x$origin, levels = c("Combined", "NOR", "HOR"))
  x = x[order(x$Class, x$Type, x$LH_type, x$origin),]

  colnames(x)[which(colnames(x) == "LH_type")] = "Mig. Type"
  colnames(x)[which(colnames(x) == "Origin")] = "Origin"

  return(x)
}
```

```{r make_ratios_table fn}
# function to make the table
make_ratios_table = function(ratios, q) {

  x = ratios
  table_x = x
  table_x[,5:9] = round(table_x[,5:9],2)
  table_x[is.na(table_x)] = ""

  # select the correct color rule
  if (q %in% c("mean", "cv")) {
    assign_color = assign_color_1
  } else {
    assign_color = assign_color_0
  }

  kable(table_x, "html", digits = 2, row.names = FALSE, align = "llllccccc") %>%
    kable_styling(full_width = FALSE, bootstrap_options = c("condensed")) %>%
    column_spec(5, background = assign_color(x[,5])) %>%
    column_spec(6, background = assign_color(x[,6])) %>%
    column_spec(7, background = assign_color(x[,7])) %>%
    column_spec(8, background = assign_color(x[,8])) %>%
    column_spec(9, background = assign_color(x[,9])) %>%
    collapse_rows(1:4)
}
```
